Choosing TTC x Provides the Most Time Stability during Rush Hours

### 2024-09-22 outputs/paper/paper.qmd
moved data-filtering steps (initially in result section of paper) to cleaning script due to excess code chunks in paper (unorganized)

```{r}
#| include: false
#| warning: false
#| message: false
# load librarys and datasets
library(tidyverse)
library(dplyr)
library(lubridate)
library(zoo)
library(knitr)
library(ggplot2)

bus_delay <- read.csv(here::here("inputs/data/cleaned_data/bus_delay.csv"))
streetcar_delay <- read.csv(here::here("inputs/data/cleaned_data/streetcar_delay.csv"))
subway_delay <- read.csv(here::here("inputs/data/cleaned_data/subway_delay.csv"))
subway_delay_code <- read.csv(here::here("inputs/data/cleaned_data/subway_delay_code.csv"))

@@

# Generate the bar plot
ggplot(delay_count, aes(x = service_type, y = num_delay, fill = service_type)) +
  geom_bar(stat = "identity") + 
  labs(title = "Number of Delays by Service Type", x = "Service Type", y = "Number of Delays", fill = "Service Type") +  # Label the axes
  theme_minimal()  # Apply a minimal theme
```

## Most Delayed during Weekday
Define Rush Hours

@@

```{r}
#| echo: false
#| message: false

#filter out delays that occurred on weekdays
weekday_bus_delay <- bus_delay |>
  filter(day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
weekday_streetcar_delay <- streetcar_delay |>
  filter(day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
weekday_subway_delay <- subway_delay |>
  filter(day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))

#filter out delays that occurred on weekends
weekend_bus_delay <- bus_delay |>
  filter(day %in% c("Saturday", "Sunday"))
weekend_streetcar_delay <- streetcar_delay |>
  filter(day %in% c("Saturday", "Sunday"))
weekend_subway_delay <- subway_delay |>
  filter(day %in% c("Saturday", "Sunday"))
```

```{r}
#| label: fig-weekday
#| fig-cap: Visualization to show the most delayed services on weekdays rush hours and off-peak hours
#| echo: false
#| message: false

@@

```{r}
#| label: fig-weekend
#| fig-cap: Visualization to show the most delayed services on weekdays rush hours
#| echo: false
#| message: false

## Most Delayed during Weekends

```{r}
#| label: fig-informed
#| fig-cap: Visualization to show correlation between number of callers given information on homelessness prevention (represented by the dark green line) and Homeless Death Counts (represented by the blue line)
#| echo: false
#| message: false

>>>>

```{r}
#| include: false
#| warning: false
#| message: false
# load librarys and datasets
library(tidyverse)
library(dplyr)
library(lubridate)
library(zoo)
library(knitr)
library(ggplot2)
library(hms)

bus_delay <- read.csv(here::here("inputs/data/cleaned_data/bus/bus_delay.csv"))
streetcar_delay <- read.csv(here::here("inputs/data/cleaned_data/streetcar/streetcar_delay.csv"))
subway_delay <- read.csv(here::here("inputs/data/cleaned_data/subway/subway_delay.csv"))
subway_delay_code <- read.csv(here::here("inputs/data/cleaned_data/subway_delay_code.csv"))

@@

# Generate the bar plot
ggplot(delay_count, aes(x = service_type, y = num_delay, fill = service_type)) +
  geom_bar(stat = "identity") + 
  geom_text(aes(label = paste0(num_delay, " delayed")), 
          vjust = 1, color = "white", size = 3.5, position = position_stack(vjust = 0.55)) +
  labs(title = "Number of Delays by Service Type", x = "Service Type", y = "Number of Delays", fill = "Service Type") +  # Label the axes
  theme_minimal()  # Apply a minimal theme
```

## Most Delayed during Weekday
Define Rush Hours:
  Rush Hours: 
  •	Morning Rush-Hour: Approximately 7:00 AM to 9:00 AM
	•	Evening Rush-Hour: Approximately 4:00 PM to 7:00 PM

@@

```{r}
#| echo: false
#| message: false
weekday_bus_delay <- read.csv(here::here("inputs/data/cleaned_data/bus/bus_delay.csv"))
weekday_streetcar_delay <- read.csv(here::here("inputs/data/cleaned_data/streetcar/streetcar_delay.csv"))
weekday_subway_delay <- read.csv(here::here("inputs/data/cleaned_data/subway/subway_delay.csv"))
```

```{r}
#| label: fig-weekday
#| fig-cap: Visualization to show the total delayed time during weekdays in 2023
#| echo: false
#| message: false

service_type <- c("bus", "streetcar", "subway")
sum_delay_time <- c(sum(weekday_bus_delay$min_delay), sum(weekday_streetcar_delay$min_delay), sum(weekday_subway_delay$min_delay))

total_delay_time <- data.frame(service_type, as.numeric(sum_delay_time))

# Generate the bar plot
ggplot(total_delay_time, aes(x = service_type, y = sum_delay_time, fill = service_type)) +
  geom_bar(stat = "identity") + 
  # Add labels for total hours
  geom_text(aes(label = paste0(round(sum_delay_time / 60, 2), " hours")), 
            vjust = 1, color = "white", size = 3.5, position = position_stack(vjust = 0.55)) +  
  labs(title = "Total Minutes Delayed in 2023 during Weekdays", 
       x = "Service Type", y = "Total Delayed (min)", fill = "Service Type") +  # Label the axes
  theme_minimal()

@@

## Most Delayed during Rush-Hours & Off-Peak Hours

```{r}
#| label: fig-rush-off
#| fig-cap: Horizontal Barplot used to visualize the longest average delay time for different TTC services during Rush-Hours and Off-Peak Hours
#| echo: false
#| message: false
